News Across Languages - Cross-Lingual Document Similarity and Event Tracking

### 2015-09-04 rupnik2015.tex
related work

There are four main families of approaches to cross-lingual similarity.

\textbf{Translation and dictionary based}. The most obvious way to compare documents written in different languages is to use machine translation and perform monolingual similarity, see  \cite{multilingualBook}\cite{plagiarism} for several variations of translation based approaches. One can use free tools such as Moses \shortcite{moses} or translation services, such as Google Translate\footnote{https://translate.google.com/}. There are two issues with such approaches: they solve a harder problem than needs to be solved and they are less robust to training resource quality - large sets of translated sentences are typically needed. Training Moses for languages with scarce linguistic resources is thus problematic. The issue with using online services such as Google Translate is that the APIs are limited and not free. The operation efficiency and cost requirements make translation-based approaches less suited for our system. CL-VSM \cite{plagiarism} and the approach presented in \cite{pouliquen2008story} compare documents by using dictionaries, such as EuroVoc \cite{eurovoc}. The generality of such approaches is limited by the quality of available linguistic resources, which may be scarce or non-existent for certain language pairs.

>>>>

There are five main families of approaches to cross-lingual similarity.

\textbf{Translation and dictionary based}. The most obvious way to compare documents written in different languages is to use machine translation and perform monolingual similarity, see  \cite{multilingualBook}\cite{plagiarism} for several variations of translation based approaches. One can use free tools such as Moses \shortcite{moses} or translation services, such as Google Translate\footnote{https://translate.google.com/}. There are two issues with such approaches: they solve a harder problem than needs to be solved and they are less robust to training resource quality - large sets of translated sentences are typically needed. Training Moses for languages with scarce linguistic resources is thus problematic. The issue with using online services such as Google Translate is that the APIs are limited and not free. The operation efficiency and cost requirements make translation-based approaches less suited for our system. Closely related are works Cross-Lingual Vector Space Model (CL-VSM) \cite{plagiarism} and the approach presented in \cite{pouliquen2008story} which both compare documents by using dictionaries, which in both cases are EuroVoc dictionaries \cite{eurovoc}. The generality of such approaches is limited by the quality of available linguistic resources, which may be scarce or non-existent for certain language pairs.
